---
title: "wikimapR, an R package for importing Wikimapia data as Simple Features via API"
keywords: "wikimapia, API, sf, Simple Features, R"
output: github_document
# output:
#   rmarkdown::html_vignette:
#     self_contained: no
# 
#   md_document:
#     variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = TRUE,
  message = TRUE,
  width = 120,
  comment = "#>",
  fig.retina = 2,
  fig.path = "man/figures/README-"
)
```


[![DOI](https://zenodo.org/badge/161494897.svg)](https://zenodo.org/badge/latestdoi/161494897)


`wikimapR` is an R package for accessing the raw vector data from Wikimapia via official [Wikimapia API](http://wikimapia.org/api). Map data is returned as
[Simple Features (`sf`)](https://cran.r-project.org/package=sf) objects with some of the object details included as nested `lists`.

**This package is at a VERY alpha stage. Provided 'as is'. Use with caution.**

You may also want to try [a similar package for Python](https://github.com/plandex/wikimapia-api-py).


### Installation

To install:
```{r install_package, eval=FALSE}
# Install the development version
# install.packages("devtools")
devtools::install_github("e-kotov/wikimapR")
```

To load the package and check the version:
```{r enable_package, eval=TRUE}
library(wikimapR)
packageVersion("wikimapR")
```


Load additional packages:
```{r}
library(sf)
library(magrittr)
library(dplyr)
library(purrr)
```


### Usage

#### Choose a bbox

Have a look at [https://boundingbox.klokantech.com](https://boundingbox.klokantech.com) and choose a bounding box. For example: `49.990525,53.100579,50.37642,53.334931` for Samara and it's surroundings.

```{r set_bbox, eval=TRUE}
bbox <- c(49.990525,53.100579,50.37642,53.334931) # Samara
```

#### Get the number of objects in this bounding box

Use [`box`](http://wikimapia.org/api#oldbox) API function to get the number of features in your area of interest.

```{r estimate_n, eval=TRUE}
wm <- wm_get_from_bbox(x = bbox, get_location = FALSE, meta_only = TRUE)
# wm <- wm_get_from_bbox(x = bbox, get_location = F, wm_api_key = "XXXXXXX", object_count_only = T) # use with your own API key to perform more frequent requests and avoid the warning message
wm$found
```

Now we know how many objects we have in the bounding box.

There are `page` and `count` parameters in the Wikimapia API, but you cannot request objects from a bounding box beyond 10 000 no matter how high you set the `page` parameter. So you have to subdivide the large bounding box that you have into smaller bounding boxes with a maximum of 10 000 objects each.

#### Subdivide the bounding box into smaller ones

`subdivide_bbox()` subdivides a large bounding box into smaller ones and returns `sf polygons`, or `bbox` objects or both. It is good for large areas with defaults tuned to cities like Moscow.

```{r subdivide_bbox, eval=TRUE}
small_bboxes <- subdivide_bbox(x = bbox, bbox_cell_size = 0.1, return_bbox_or_sf = "both")
plot(small_bboxes$sf$geometry)
str(small_bboxes, max.level = 1)
```

In this example the bounding box cell size is in degrees (need to somehow fix that to work with meters across the globe). Default of 0.045 degrees is reasonably large, roughly equivalent to 2845x5010 meters. It has proved to fit < 10 000 objects per cell in Moscow, where the density of objects is quite high. For less object-dense cities you may be able to get away with larger grid of bounding boxes.

This is not ideal and it would be more convenient to create this subdivision using precise metric, but this will do fine for now.

#### Check then number of objects per cell

Just to be sure that every bounding box that we generated has <= 10 000 objects, let us query all the bounding boxes. For the current example with `r length(small_bboxes$bbox)` it will take about `r length(small_bboxes$bbox)/2` minutes, as with "example" API key the cool-down is about 30 seconds.

```{r estimate_subdivided_n, cache=TRUE, message=FALSE, warning=FALSE}
objects_in_bboxes <- list()
for( i in 1:length(small_bboxes$bbox)){
  Sys.sleep(30) # wait for API required interval, sleep goes before the function itself for the map to correctly collect values in the list
    objects_in_bboxes[[i]] <- wm_get_from_bbox(x = small_bboxes$bbox[[i]], get_location = FALSE) # get objects in every bbox, but no need to get the location for now
}
```

Now we look at the histogram of the number of objects in the small bounding boxes and the maximum value:

```{r b_per_bbox, eval=TRUE}
n_by_bbox <- objects_in_bboxes %>% purrr::map_int(~ .x$meta$found) # extract the number of found objects for every bounding box
max(n_by_bbox)
hist(n_by_bbox)
```

Since the maximum is well within 10 000, we can proceed to collect the objects IDs and then download attributes for individual features. We have to download the detailed objects features one-by-one as [`box`](http://wikimapia.org/api#oldbox) API only returns object ID, name and URL. So the strategy for getting all object details is to make a list of object IDs using [`box`](http://wikimapia.org/api#oldbox) API and then use [`place.getbyid`](http://wikimapia.org/api#placegetbyid).


#### Create a list of IDs to fetch

```{r id_list, eval=TRUE, cache=TRUE}
id_list <- objects_in_bboxes %>%
  purrr::map(~ .x$df$id) %>% # pull object IDs from individual bbox query results data.frames
  do.call(c, .) # bind together

str(id_list)
head(id_list)
```

Now we have a list of `r length(id_list)` object IDs that we want to get the details for.

#### Get detailed data for Wikimapia objects

Let's take just 3 first objects for this example. It will take 1.5 minutes to fetch them with all the details using the example API.

```{r get_indiv_objects, cache=TRUE, message=FALSE, warning=FALSE}
short_list <- id_list[1:3]
wm_objects <- wm_get_by_id(x = short_list, wm_api_key = "example")
```


#### We have the objects!

```{r obj_str}
str(wm_objects, max.level = 1, nchar.max = 50)
```


##### We can now plot them

```{r obj_plot}
plot(wm_objects$geometry)
```

##### We can explore the attributes

```{r obj_columns}
names(wm_objects)
```

##### Most of the details are currently in a nested list for every object

You can use [purrr](https://github.com/tidyverse/purrr) and/or [rlist](https://github.com/renkun-ken/rlist) packages to pull any of the details from these nested lists.

```{r obj_details}
str(wm_objects$details[[1]], max.level = 1)
```


### To-Do List

- Rewrite `subdivide_bbox()` to accept metric values for bbox dimensions (not critical, but may be useful for other projects)

- Create a few more helper functions to abstract the user from the calls to `purrr` for simple things like getting the number of found objects, or for getting number of objects per bounding box (see examples above in the Usage section)

- Create a hidden environment variable for storing API key and using it automatically in the package functions

- Implement the rest of the Wikimapia API functions

- Fix bugs if any

- Make code more robust

- Write unit tests

- Your suggestions are welcome via GitHub issues for this package

- Submit the package to CRAN someday..?

### Citation

```{r}
citation ("wikimapR")
```

### License

The MIT License (MIT) + License File

Copyright © 2018 Egor Kotov

Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the “Software”), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED “AS IS”, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.

